Imputation of Incomplete Multilevel Data with \pkg{mice}

### 2021-11-04 Manuscript/Manuscript.tex
Add data exploration popularity dataset

\texttt{impact} is traumatic brain injury data with patients clustered
in studies, \(n_{\text{participants}} = 11022\) and
\(n_{\text{clusters}} = 15\), on the following 11 variables: *
\texttt{name} Name of the study, * \texttt{type} Type of study (RCT:
randomized controlled trial, OBS: observational cohort), * \texttt{age}
Age of the patient, * \texttt{motor\_score} Glasgow Coma Scale motor
score, * \texttt{pupil} Pupillary reactivity, * \texttt{ct} Marshall
Computerized Tomography classification, * \texttt{hypox} Hypoxia (0=no,
1=yes), * \texttt{hypots} Hypotension (0=no, 1=yes), * \texttt{tsah}
Traumatic subarachnoid hemorrhage (0=no, 1=yes), * \texttt{edh} Epidural
hematoma (0=no, 1=yes), * \texttt{mort} 6-month mortality (0=alive,
1=dead).

\begin{CodeChunk}
\begin{CodeInput}
R> # load data
R> data("impact")
R> # # descriptive statistics
R> # by(impact, impact$name, summary) 
R> # psych::describe(impact)[,c(2:5,8:9)]
R> # missingness
R> md_pat(impact)
\end{CodeInput}
\begin{CodeOutput}
 /\     /\
{  `---'  }
{  O   O  }
==>  V <==  No need for mice. This data set is completely observed.
 \  \|/  /
  `-----'
\end{CodeOutput}

@@

\texttt{popNCR} is a simulated dataset with pupils clustered in classes,
\(n_{\text{participants}} = 2000\), \(n_{\text{clusters}} = 100\), on 7
variables: * \texttt{pupil} Pupil number within class, * \texttt{class}
Class number, * \texttt{extrav} Pupil extraversion, * \texttt{sex} Pupil
gender, * \texttt{texp} Teacher experience (years), * \texttt{popular}
Pupil popularity, * \texttt{popteach} Teacher popularity.

\begin{CodeChunk}
\begin{CodeInput}
R> # load data
R> pop <- readRDS("../Data/popNCR.RDS")
R> # missingness
R> md_pat(pop)
\end{CodeInput}

\begin{center}\includegraphics{Manuscript_files/figure-latex/pop-1} \end{center}

@@

\begin{itemize}
\item
  Which models will we discuss? We'll build the model to grow in
  complexity. The final model is the most complex but also the most
  versatile.
\item
  Note on model complexity: Typically, we should at least use random
  intercepts, but often random slopes as well. Ideally we impute with
  random everything and heteroscedastic errors: most generic method (no
  worry about congeniality, but don't mention the term) -\textgreater{}
  Refer to other papers for background, we'll focus just on the software
  implementation of the situations mentioned there. Sometimes there's
  little reason to assume some variable is affected by heterogeneity.
  -\textgreater{} Refer to Meng, Vincent, and a paper by Grund on
  congeniality and random slopes.
\item
  Step 0: study as a predictor, AKA multilevel imputation for dummies.
  Doesn't work for syst missing.
\end{itemize}

\hypertarget{conditional-models}{%
\subsection{Conditional models}\label{conditional-models}}

\begin{itemize}
\item
  How to define the imputation model(s) in \texttt{mice}?
\item
  What do the different implementations look like?
\item
  Step 1: Intercept
\item
  Step 2: Slope
\item
  Step 3: Residuals
\item
  Heckman model for MNAR
\end{itemize}

>>>>

\begin{itemize}
\item
  What does the data look like? \texttt{impact} is traumatic brain
  injury data with patients clustered in studies,
  \(n_{\text{participants}} = 11022\) and \(n_{\text{clusters}} = 15\),
  on the following 11 variables:

  \begin{itemize}
  \tightlist
  \item
    \texttt{name} Name of the study,
  \item
    \texttt{type} Type of study (RCT: randomized controlled trial, OBS:
    observational cohort),
  \item
    \texttt{age} Age of the patient,
  \item
    \texttt{motor\_score} Glasgow Coma Scale motor score,
  \item
    \texttt{pupil} Pupillary reactivity,
  \item
    \texttt{ct} Marshall Computerized Tomography classification,
  \item
    \texttt{hypox} Hypoxia (0=no, 1=yes),
  \item
    \texttt{hypots} Hypotension (0=no, 1=yes),
  \item
    \texttt{tsah} Traumatic subarachnoid hemorrhage (0=no, 1=yes),
  \item
    \texttt{edh} Epidural hematoma (0=no, 1=yes),
  \item
    \texttt{mort} 6-month mortality (0=alive, 1=dead).
  \end{itemize}
\end{itemize}

\begin{CodeChunk}
\begin{CodeOutput}
 /\     /\
{  `---'  }
{  O   O  }
==>  V <==  No need for mice. This data set is completely observed.
 \  \|/  /
  `-----'
\end{CodeOutput}

@@

\begin{itemize}
\item
  What does the data look like? \texttt{popNCR} is a simulated dataset
  with pupils clustered in classes, \(n_{\text{participants}} = 2000\),
  \(n_{\text{clusters}} = 100\), on 7 variables:

  \begin{itemize}
  \tightlist
  \item
    \texttt{pupil} Pupil number within class,
  \item
    \texttt{class} Class number,
  \item
    \texttt{extrav} Pupil extraversion,
  \item
    \texttt{sex} Pupil gender,
  \item
    \texttt{texp} Teacher experience (years),
  \item
    \texttt{popular} Pupil popularity,
  \item
    \texttt{popteach} Teacher popularity.
  \end{itemize}
\item
  What are the ICCs? For \texttt{popular} the ICC is 0.33. For
  \texttt{popteach} it is 0.31.
\item
  What does the missingness look like? Induced MAR/MNAR missingness.
  Missing data pattern:
\end{itemize}

\begin{center}\includegraphics{Manuscript_files/figure-latex/pop_pat-1} \end{center}

\begin{itemize}
\item
  Does the missing data of popular depend on popteach? One could for
  example check this by making a histogram of popteach separately for
  the pupils with known popularity and missing popularity. We do see
  that the histogram for the missing popular (TRUE) is further to the
  right than the histogram for observed popular (FALSE). This would
  indicate a right-tailed MAR missingness. In fact this is exactly what
  happens, because we created the missingness in these data ourselves.
  But we can make it observable by examining the relations between the
  missingness in popular and the observed data in popteach.
\item
  Does the missingness in teacher popularity depend on pupil popularity?
  Yes: there is a dependency. The relation seems to be right-tailed.
\end{itemize}

\begin{center}\includegraphics{Manuscript_files/figure-latex/pop_dist-1} \end{center}

\begin{center}\includegraphics{Manuscript_files/figure-latex/pop_dist-2} \end{center}

@@

\begin{itemize}
\item
  Which models will we discuss? We'll build the model to grow in
  complexity. The final model is the most complex but also the most
  versatile.
\item
  Note on model complexity: Typically, we should at least use random
  intercepts, but often random slopes as well. Ideally we impute with
  random everything and heteroscedastic errors: most generic method (no
  worry about congeniality, but don't mention the term) -\textgreater{}
  Refer to other papers for background, we'll focus just on the software
  implementation of the situations mentioned there. Sometimes there's
  little reason to assume some variable is affected by heterogeneity.
  -\textgreater{} Refer to Meng, Vincent, and a paper by Grund on
  congeniality and random slopes.
\item
  Step 0: As predictor
\item
  Step 1: Random intercepts
\item
  Step 2: Random slopes
\item
  Step 3: Residuals
\item
  Heckman model for MNAR
\item
  What do the different implementations look like? How to define the
  imputation model(s) in \texttt{mice}?
\end{itemize}

\hypertarget{step-0}{%
\subsection{Step 0}\label{step-0}}

\begin{itemize}
\item
  AKA multilevel imputation for dummies.
\item
  Doesn't work for systematic missingness.
\end{itemize}

\hypertarget{step-1-3}{%
\subsection{Step 1-3}\label{step-1-3}}

\begin{itemize}
\tightlist
\item
  TODO: fill in.
\end{itemize}
